---
title: "facebook anonymized data"
author: "~~~~"
date: "April 18, 2018"
output:
  html_document:
    df_print: paged
---

# the package we need is igraph( a powerful graphing package)
#                        datatable(a package to help read data)
```{r library}
source("../lib/functions.R")
packages.used <- c("data.table","igraph","factoextra","dplyr","ggplot2","mixedMem","cluster")

# check packages that need to be installed.
packages.needed <- setdiff(packages.used, intersect(installed.packages()[,1], packages.used))

# install additional packages
if(length(packages.needed) > 0) {
  install.packages(packages.needed, dependencies = TRUE, repos = 'http://cran.us.r-project.org')
}

library("data.table");library(igraph);library(reshape2);library(factoextra)
library("dplyr");library("ggplot2");library(mixedMem);library(cluster)
```

# This dataset consists of 'circles' (or 'friends lists') from Facebook. Facebook data was collected from survey participants using this Facebook app.

# Facebook data has been anonymized by replacing the Facebook-internal ids for each user with a new value.

# the FROM and TO are Undirected! (friends relation are bidirectional)
```{r data}
data<-data.table(read.table("../data/facebook_combined.txt"))
colnames(data) <-c("FROM", "TO")
# starting from 1
data<-data+1
dim(data)
data[1:5]
```


# use igraph to plot the data, igraph is powerful, we can edit the graph in a way we like
# to see how graph looks like
```{r plot example1}
example<-data
example<-as.matrix(example)
gr <-graph_from_edgelist(example,directed = FALSE)
gr <- delete_vertices(gr,which(degree(gr)==0))   
#E(gr)$color <- ("blue")
V(gr)$color <- ("red")
V(gr)$size <- 4

# take time to plot!!!
graph1<-plot(gr, vertex.color = adjustcolor("SkyBlue2", alpha.f = .5), 
      vertex.label.color = adjustcolor("black", 0.2))

```

##1. Hierarchical clustering with Agglomerative-Euclidean-single-linkage clustering approach:

# First we transfer the data into a matrix form:
```{r transfer to matrix}
matrixdata<-shan_zhong_matrix_data_transform(data)
dim(matrixdata)
matrixdata[1:12,1:12]
```

#Then we calculate the distance for each user

#Let's start to look at user 1, one of the 10 people manually identified all the circles to which his friends belonged

#User1 are friends of user2 to user348
#The computation time do not vary much

```{r compare time for distance}
# since the computation time do not make much difference, we choose euclidean method
system.time(dist(matrixdata[1:348,1:348],method="manhattan"))
system.time(dist(matrixdata[1:348,1:348],method="euclidean"))
system.time(dist(matrixdata[1:348,1:348],method="maximum"))
```

#Now we calculate the distance for each circle, there are many method to define the distance between two circles:

#Single : Minimum over all pairs of users these two circle.

#Complete : Maximum over all pairs of users these two circle.

#Average : Average over all pairs of users these two circle.

```{r compare for linkage}
# since the computation time do not make much difference, we choose euclidean method

distance<-dist(matrixdata[1:348,1:348],method="euclidean")

single<-hclust(distance,method="single")
complete<-hclust(distance,method="complete")
average<-hclust(distance,method="average")

plot(single)
plot(complete)
plot(average)

# Optimal number of circles, for hierarchical modeling(elbow method, independent of previous methods we choose):

fviz_nbclust(matrixdata[1:348,1:348],FUN=hcut,method= "wss",k.max= 25)

fviz_nbclust(matrixdata[1:348,1:348],FUN=hcut,method= "silhouette",k.max= 25)

# gap_stat<-clusGap(matrixdata[1:346,1:346], FUN = hcut, nstart = 25, K.max = 25, B=50)
# fviz_nbclust(gap_stat)

```

## model 2, structural cohesion, use package from igraph
# structural cohesion for the circle of user 1, all friends of user 1 are shown in the graph!
```{r plot Structural cohesion}
user1<-data[FROM %in% 1:348 & TO %in% 1:348]
user1<-as.matrix(user1)
gr <-graph_from_edgelist(user1, directed = FALSE)

gr <- delete_vertices(gr,which(degree(gr)==0))                      
                      
grBlocks <- cohesive_blocks(gr)

E(gr)$color <- ("blue")
V(gr)$color <- ("red")
V(gr)$size <- 4

# take time to plot!!!
plot(grBlocks, gr)
```


# plot the matrix data
```{r plot data matrix}
data1<-shan_zhong_data_forggplot2(data[FROM %in% 1:348 & TO %in% 1:348])

plot<-ggplot(data1,aes(x =FROM, y = TO))+
  geom_raster()+theme_bw()+coord_flip()+
  scale_x_reverse()+
  theme(aspect.ratio = 1, legend.position="none")
#plot
```

### Mix membership model:

```{r mixmembership}

# we choose 9 
K<-9

initial<-shan_zhong_data_formixMem(matrixdata[1:348,1:348],K)

computeELBO(initial)

st = proc.time()
#printStatus 1 indicates that status updates will be printed
# printMod 25 indicates that status updates will only be printed ever 25th step

out <- mmVarFit(initial, printStatus = 1, printMod = 25)

end <- proc.time()

computeELBO(out)

end-st

```

### visualize membership model
```{r plot Mix membership}
load("../output/out.Rdata")
 plot_data<-shan_zhong_mixMemplot(out$theta,data[FROM %in% 1:348 & TO %in% 1:348])
 plot_data2<-shan_zhong_mixMemplot_order(out$theta,data[FROM %in% 1:348 & TO %in% 1:348])
 
ggplot(plot_data, aes(x = from, y = to, fill = group)) +
  geom_raster() +coord_flip()+ theme_bw() +
  theme(legend.position = "none")

ggplot(plot_data2, aes(x = from, y = to, fill = group)) +
  geom_raster() +coord_flip()+ theme_bw() +
  theme(legend.position = "none")
 
```